---
title: "Example `reactflow` of a `targets` Pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{example_targets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The [`targets`](https://books.ropensci.org/targets/) R package is a wonderful tool for reproducible data science workflows.
To visualize a targets pipeline, there is the default `tar_visnetwork()` function, which creates a network visualization of the pipeline.

As this information is by definition a graph, `reactflow` can be used to visualize and style the pipeline in a more interactive way.

Note we won't be covering the `targets` package in this vignette, the code to create the targets pipeline etc is included in `reactflow` in `data-raw/example_targets.R` on Github.

This `reactflow` package provides two sample datasets, `targets_meta` (output of `targets::tar_meta()`) and `targets_network` (output of `targets::tar_network()`).

To prepare the data for the `reactflow` visualization, we need to convert the `targets_network` data to a format that `reactflow` can understand.
That is, we need to convert the data.frame to a list of nodes and a list of edges.

```{r setup}
library(reactflow)
library(bsicons)
tar <- targets_network

# merge meta information to identify files
tar$vertices <- merge(tar$vertices, targets_meta[, c("name", "format")], 
                      by.x = "name", by.y = "name")

icons <- list(
  file = bs_icon("file-earmark-fill", size = 15),
  stem = bs_icon("type", size = 15),
  "function" = bs_icon("braces", size = 15)
)

nodes <- lapply(seq_len(nrow(tar$vertices)), \(i) {
  el <- tar$vertices[i, ]
  
  name <- el$name
  icon <- icons[[ifelse(!is.na(el$format) & el$format == "file", "file", el$type)]]
  time <- ifelse(el$status == "uptodate", 
                 ifelse(is.na(el$seconds), "NA", paste0(el$seconds, "s")),
                 "NA")
  
  
  html <- as.character(
    htmltools::div(
      style = "display: flex; flex-direction: row; align-items: center;",
      class = el$status,
      htmltools::div(icon, style = "margin-right: 10px;"),
      htmltools::div(
        htmltools::div(
          style = "font-size: 1.1em; font-weight: bold; color: #999999",
          name
        ),
        htmltools::div(
          style = "font-size: 0.8em; font-weight: 600; color: darkgray;",
          paste0("Time: ", time)
        )
      )
    )
  )
  list(
    id = el$name,
    type = "html",
    data = list(html = html,
                status = el$status, # for mini-map later on
                handle = list(direction = "LR"))
  )
})


edges <- apply(tar$edges, 1, \(x) {
  list(
    id = paste(x[["from"]], x[["to"]], sep = "_"),
    source = x[["from"]],
    target = x[["to"]]
  )
})
```

Note we also need to set some general CSS to make the nodes look a bit nicer

```{css}
.react-flow__node.react-flow__node-html {
  padding: 10px;
  border-radius: 15px;
  border: 4px solid gray;
  width: 150px;
  font-size: 12px;
  text-align: center;
  background-color: var(--xy-node-background-color, var(--xy-node-background-color-default));
}

.react-flow__node.react-flow__node-html:has(.outdated) {
  border-color: lightcoral;
  color: lightcoral;
}
.react-flow__node.react-flow__node-html:has(.uptodate) {
  border-color: mediumseagreen;
  color: mediumseagreen;
}

/* style the handles */
.react-flow__handle{
  width: 5px;
  height: 5px;
  border-radius: 100px;
}

/* make edges a bit thicker */
.react-flow__edge > path {
    stroke: gray;
    stroke-width: 3px;
}
```


Last but not least, we can create our reactflow like this:

```{r}
reactflow(
  nodes = nodes,
  edges = edges,
  mini_map(
    pannable = TRUE,
    zoomable = TRUE,
    nodeColor = JS("
    function nodeColor(node) {
      switch (node.data.status) {
        case 'uptodate':
          return 'mediumseagreen';
        default:
          return 'lightcoral';
      }
    }")
  ),
  controls(),
  background(),
  
  width = "100%",
  allow_edge_connection = FALSE,
  use_dagre = TRUE,
  dagre_direction = "LR",
  dagre_config = list(nodeWidth = 200, nodeHeight = 50)
)
```
