---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

  
```{r}
library(reactflow)
```


# Basic Usage

To create a basic flow using `reactflow`, you need to define nodes and edges and call the `reactflow` function with these arguments.

Note to label a node, you provide a `label` argument in the `data` list, whereas the `label` is a direct argument to an edge.

```{r}
nodes <- list(
  list(id = "1", position = list(x = 0, y = 0), data = list(label = "Node 1")),
  list(id = "2", position = list(x = 100, y = 100), data = list(label = "Node 2")),
  list(id = "3", position = list(x = 0, y = 200), data = list(label = "Node 3"))
)

edges <- list(
  list(id = "e1-2", source = "1", target = "2", label = "Edge 1"),
  list(id = "e2-3", source = "2", target = "3"),
  list(id = "e1-3", source = "1", target = "3")
)
reactflow(
  nodes,
  edges
)
```

You can change the width of the output widget directly in the call to [`reactflow()`]:

```{r}
reactflow(
  nodes,
  edges,
  width = "250px"
)
```

By default, you can use connect new edges by dragging from one node to another.
To disable this, you can set

```{r}
reactflow(
  nodes,
  edges,
  allow_edge_connection = FALSE
)
```



## UI Elements

The `reactflow` package provides some easy to use UI elements to customize the usage and display of a flow.
Namely [`background()`], [`controls()`], [`mini_map()`],

```{r}
reactflow(
  nodes,
  edges,
  background = background(),
  controls = controls(),
  mini_map = mini_map()
)
```

## Background

All arguments passed to `background()` are passed to the underlying `react-flow` component.
The official `background` documentation can be found here: <https://reactflow.dev/api-reference/components/background>.

```{r}

nodes <- list(
  list(id = "1", position = list(x = 0, y = 0), data = list(label = "Node 1")),
  list(id = "2", position = list(x = 100, y = 100), data = list(label = "Node 2")),
  list(id = "3", position = list(x = 0, y = 200), data = list(label = "Node 3"))
)

edges <- list(
  list(id = "e1-2", source = "1", target = "2", label = "Edge 1"),
  list(id = "e2-3", source = "2", target = "3"),
  list(id = "e1-3", source = "1", target = "3")
)

reactflow(
  nodes,
  edges,
  background = background()
)
```

An example using the extra arguments can be found below

```{r}
reactflow(
  nodes,
  edges,
  background = background(
    color = "beige",
    variant = "lines",
    gap = 50,
    lineWidth = 2
  )
)
```

### Multiple Backgrounds

You can also add multiple backgrounds to the flow by passing multiple `background()` elements.

```{R}
reactflow(
  nodes,
  edges,
  background(
    id = 1,
    color = "beige",
    variant = "lines",
    gap = 40,
    lineWidth = 3
  ),
  background(
    id = 2,
    size = 4,
    offset = 3,
    color = "lightgray",
    variant = "dots"
  )
)
```

## Controls

All arguments passed to `controls()` are passed to the underlying `react-flow` component.
The official `controls` documentation can be found here: <https://reactflow.dev/api-reference/components/controls>.


```{r}
nodes <- list(
  list(id = "1", position = list(x = 0, y = 0), data = list(label = "Node 1")),
  list(id = "2", position = list(x = 100, y = 100), data = list(label = "Node 2")),
  list(id = "3", position = list(x = 0, y = 200), data = list(label = "Node 3"))
)

edges <- list(
  list(id = "e1-2", source = "1", target = "2", label = "Edge 1"),
  list(id = "e2-3", source = "2", target = "3"),
  list(id = "e1-3", source = "1", target = "3")
)

reactflow(
  nodes,
  edges,
  controls = controls()
)
```

An example using the extra arguments can be found below

```{r}
reactflow(
  nodes,
  edges,
  controls = controls(
    position = "top-right",
    orientation = "horizontal",
    showZoom = TRUE,
    showFitView = FALSE,
    showIntereactive = TRUE
  )
)
```


## Mini Map

All arguments passed to `mini_map()` are passed to the underlying `react-flow` component.
The official `miniMap` documentation can be found here: <https://reactflow.dev/api-reference/components/minimap>.

```{r}
nodes <- list(
  list(id = "1", position = list(x = 0, y = 0), data = list(label = "Node 1")),
  list(id = "2", position = list(x = 100, y = 100), data = list(label = "Node 2")),
  list(id = "3", position = list(x = 0, y = 200), data = list(label = "Node 3"))
)

edges <- list(
  list(id = "e1-2", source = "1", target = "2", label = "Edge 1"),
  list(id = "e2-3", source = "2", target = "3"),
  list(id = "e1-3", source = "1", target = "3")
)

reactflow(
  nodes,
  edges,
  mini_map = mini_map(pannable = TRUE, zoomable = TRUE)
)
```


# Styling

Each component can be styled individually or using CSS.
Eg setting the following CSS class, will make all nodes green.

```css
.react-flow__node {
  background-color: green !important;
}
```


See also `app.R` TODO


## Nodes

### Node Types

Note that nodes have different types, eg `type = "input", "output", "default", and "group"`.
See also official documentation at <https://reactflow.dev/api-reference/types/node>.

Node types of `input` do not have any ingoing connections, whereas `output`s do not have any outgoing connections.



### Direct Styling of Nodes

```{r}
nodes2 <- list(
  list(id = "1", position = list(x = 0, y = 0),
       data = list(label = "I have direct Styling"),
       style = list("background-color" = "#2196f3",
                 color = "white",
                 "border-color" = "#0964AE",
                 "border-width" = "2px",
                 "font-size" = "1.25em",
                 filter = "drop-shadow(10px 10px 5px gray)")),
  list(id = "2", position = list(x = 100, y = 100), data = list(label = "Node 2")),
  list(id = "3", position = list(x = 0, y = 200), data = list(label = "Node 3"))
)
reactflow(
  nodes = nodes2,
  edges = edges
)
```


### Using `className` and custom CSS

Using `className`s, we can assign custom CSS to the nodes in the mini-map.

```{css}
.startNode {
  background-color: lightgreen;
}
.endNode {
  background-color: coral;
}
```


```{r}
nodes2 <- list(
  list(id = "1", position = list(x = 0, y = 0), data = list(label = "Node 1"),
       className = "startNode"),
  list(id = "2", position = list(x = 100, y = 100), data = list(label = "Node 2")),
  list(id = "3", position = list(x = 0, y = 200), data = list(label = "Node 3"),
       className = "endNode")
)
reactflow(
  nodes = nodes2,
  edges = edges,
  mini_map = mini_map(),
  color_minimap = TRUE
)
```


### Styling using `htmlNodes`

The reactflow R package brings its own `html` node Type, which allows you to style a node using HTML and CSS.

To use this, you have to set the type of the node to `html` and provide the HTML code in the `html` data argument.

```{r}
nodes2 <- list(
  list(
    id = "1",
    type = "html", # this is required
    data = list(
      html = "I am a <b>HTML</b> node" # define your own HTML here
    ),
    position = list(x = 0, y = 0)
  ),
  list(
    id = "2",
    type = "html",
    data = list(
      html = as.character( # the HTML can be created using htmltools or shiny as well!
        htmltools::div(
          style = paste(
            "color: red; border: 2px solid gray; padding: 10px;",
            "border-radius: 10px; font-weight: 600;"
          ),
          
          "I am a ", htmltools::code("HTML"), "node"
        )
      )
    ),
    position = list(x = 200, y = 100)
  )
)

edges2 <- list(
  list(id = "e1-2", source = "1", target = "2", label = "Edge 1")
)

reactflow(
  nodes2,
  edges2
)
```

See also the Vignette on Fully Customized HTML Nodes.


### LR Layouts

If you are using Left-to-Right (LR) layouts, either by specifying positions manually or using `use_dagre = TRUE, dagre_direction = "LR"`, you can use a node with `data.handle.position = "LR"` to make the handle appear on the left side of a `html` node.

```{r}
nodes2 <- list(
  list(
    id = "1",
    type = "html",
    data = list(
      html = "I am a <b>HTML</b> node",
      handle = list(direction = "LR") # this is new here!
    ),
    position = list(x = 0, y = 0)
  ),
  list(
    id = "2",
    type = "html",
    data = list(
      html = as.character( # the HTML can be created using htmltools or shiny as well!
        htmltools::div(
          style = paste(
            "color: red; border: 2px solid gray; padding: 10px;",
            "border-radius: 10px; font-weight: 600;"
          ),
          
          "I am a ", htmltools::code("HTML"), "node"
        )
      ),
      handle = list(direction = "LR") # this is new here!
    ),
    position = list(x = 200, y = 0)
  )
)

edges2 <- list(
  list(id = "e1-2", source = "1", target = "2", label = "Edge 1")
)

reactflow(
  nodes2,
  edges2
)
```


## Edges

### Edge Types and Styling

ReactFlow knows different edge types, these are

- `straight`
- `step`
- `smoothstep`
- `simplebezier`

Additionally, each edge can be animated by setting `animated = TRUE`.

The full documentation can be found here:
<https://reactflow.dev/api-reference/types/edge>


```{r}
nodes2 <- list(
  list(id = "1", position = list(x = 0, y = 0), data = list(label = "Node 1")),
  list(id = "2", position = list(x = 100, y = 100), data = list(label = "Node 2")),
  list(id = "3", position = list(x = 200, y = 200), data = list(label = "Node 3"))
)

edges2 <- list(
  list(id = "e1-2", source = "1", target = "2", label = "straight",
       type = "straight"),
  list(id = "e2-3", source = "2", target = "3", label = "smoothstep + animated", 
       type = "smoothstep", animated = TRUE),
  list(id = "e1-3", source = "1", target = "3", label = "simplebezier",
       type = "simplebezier")
)

reactflow(
  nodes2,
  edges2
)
```

Note that you can slow down the animation by using the following CSS
```css
.react-flow__edge.animated path {
  animation: dashdraw 2s linear infinite;
}
```
```{css, echo=FALSE}
.slow_anim .react-flow__edge.animated path {
  animation: dashdraw 2s linear infinite;
}
```

<div class="slow_anim">

```{R, class.source="slow_anim"}
edges2 <- list(
  list(id = "e1-2", source = "1", target = "2", label = "straight",
       type = "straight", animated = TRUE),
  list(id = "e2-3", source = "2", target = "3", label = "smoothstep + animated", 
       type = "smoothstep", animated = TRUE),
  list(id = "e1-3", source = "1", target = "3", label = "simplebezier",
       type = "simplebezier", animated = TRUE)
)

reactflow(
  nodes,
  edges2
)
```

</div>


### Direct Styling

### Using `className` and custom CSS



## Mini Map

The official `miniMap` documentation can be found here: <https://reactflow.dev/api-reference/components/minimap>.
For example, the `nodeColor` argument can be used to change the color of the nodes in the mini map.

```{r}
reactflow(
  nodes,
  edges,
  mini_map = mini_map(
    pannable = TRUE,
    zoomable = TRUE,
    nodeColor = "coral",
    position = "top-right",
    maskColor = "lightgray"
  )
)
```

Using `className`s, we can assign custom CSS to the nodes in the mini-map.

```{css}
.startNode {
  background-color: lightgreen;
}
.endNode {
  background-color: coral;
}
```


```{r}
nodes2 <- list(
  list(id = "1", position = list(x = 0, y = 0), data = list(label = "Node 1"),
       type = "input"),
  list(id = "2", position = list(x = 100, y = 100), data = list(label = "Node 2")),
  list(id = "3", position = list(x = 0, y = 200), data = list(label = "Node 3"),
       type = "output")
)
reactflow(
  nodes = nodes2,
  edges = edges,
  mini_map = mini_map(
    nodeColor = JS("
    function nodeColor(node) {
      switch (node.type) {
        case 'input':
          return 'coral';
        case 'output':
          return 'lightblue';
        default:
          return 'gray';
      }
    }")
  )
)
```



## Background



## Controls

There is very little styling that can be applied to [`controls()`] directly.
However, we can apply styling using the CSS class `react-flow__controls`.


# Layouting

Normally, the layout of the nodes is done automatically by specifying the `position` arguments of the nodes.
There are however some layouting algorithms that can be used to automatically layout the nodes.
One solution is to use a tool like [`ggraph`](https://www.data-imaginist.com/posts/2017-02-06-ggraph-introduction-layouts/index.html) and its [`create_layout()`](https://ggraph.data-imaginist.com/reference/ggraph.html) function (with some transformation done to x and y).

Alternatively, you can use the built-in [`dagre`](https://github.com/dagrejs/dagre) solution.

## Automatic Layout using `dagre`

```{r}
nodes_orig <- list(
  list(id = "start", type = "input", data = list(label = "Start Node")),
  list(id = "mid", data = list(label = "Mid Node")),
  list(id = "mid2", data = list(label = "Mid Node 2")),
  list(id = "end", type = "output", data = list(label = "End Node"))
)
edges_orig <- list(
  list(source = "start", target = "mid", id = "start-mid"),
  list(source = "start", target = "mid2", id = "start-mid2"),
  list(source = "mid", target = "end", id = "mid-end"),
  list(source = "mid2", target = "end", id = "mid2-end")
)
reactflow(
  nodes = nodes_orig,
  edges = edges_orig,
  
  use_dagre = TRUE, dagre_direction = "TB",
  dagre_config = list(nodeWidth = 250, nodeHeight = 50)
)
```


# Interactivity with Shiny

```r
library(shiny)
library(bslib)
library(reactflow)

ui <- page_sidebar(
  sidebar = sidebar(
    h4("Add Node"),
    textInput("node_name", "Node Name"),
    selectInput("node_prev", "Previous Node", ""),
    selectInput("node_next", "Next Node", ""),
    actionButton("add_node", "Add Node"),
    hr(),
    h4("Selected Element"),
    verbatimTextOutput("selected_el")
  ),
  card(
    reactflowOutput("my_graph", height = "100%"),
  )
)

nodes_orig <- list(
  list(id = "start", data = list(label = "Start Node")),
  list(id = "mid", data = list(label = "Mid Node")),
  list(id = "end", data = list(label = "End Node"))
)

edges_orig <- list(
  list(source = "start", target = "mid", id = "start-mid"),
  list(source = "mid", target = "end", id = "mid-end")
)

server <- function(input, output, session) {
  output$selected_el <- renderPrint(str(input$my_graph_click))

  nodes <- reactiveVal(nodes_orig)
  edges <- reactiveVal(edges_orig)
  
  observeEvent(nodes(), {
    all_nodes <- sapply(nodes(), "[[", "id")
    updateTextInput(session, "node_name", value = "")
    updateSelectInput(session, "node_prev", choices = c("", all_nodes))
    updateSelectInput(session, "node_next", choices = c("", all_nodes))
  })
  
  
  observeEvent(input$add_node, {
    new_id <- as.character(length(nodes()) + 1)
    
    new_node <- list(id = new_id, data = list(label = input$node_name))
    
    from <- input$node_prev
    to <- input$node_next
    
    new_edge <- list(
      if (from != "") list(id = paste(from, new_id, sep = "-"), source = from, target = new_id),
      if (to != "") list(id = paste(new_id, to, sep = "-"), source = new_id, target = to)
    )
    new_edge <- new_edge[sapply(new_edge, length) > 0]
    
    nodes(c(nodes(), list(new_node)))
    edges(c(edges(), new_edge))
  })
  
  output$my_graph <- renderReactflow({
    reactflow(
      nodes = nodes(),
      edges = edges(),
      
      controls(),
      clickId = "my_graph_click",
      
      use_dagre = TRUE, dagre_direction = "TB",
      dagre_config = list(nodeWidth = 250, nodeHeight = 50),
      fitView = TRUE,
      allow_edge_connection = FALSE, # disable edge creation
      nodesDraggable = TRUE # default value, can be set to FALSE to disable dragging
    )
  })
}

shinyApp(ui, server)
```


## Styling Selected Elements

Using CSS, we can style the selected elements in the flow.
Note that it's easier to override the CSS variable for coloring the edges.
See also a list of all ReactFlow CSS variables here: <https://reactflow.dev/learn/customization/theming>

```css
.react-flow__node.selected {
  background-color: coral !important;
  color: white !important;
}
:root {
  --xy-edge-stroke-selected: coral;
}
```

```r
library(shiny)
library(bslib)
library(reactflow)

ui <- page_sidebar(
  tags$head(tags$style(
    ".react-flow__node.selected {
       background-color: coral !important;
       color: white !important;
     }
     :root {
        --xy-edge-stroke-selected: coral;
      }"
  )),
  card(
    reactflowOutput("my_graph")
  )
)

server <- function(input, output, session) {
  output$my_graph <- renderReactflow({
    reactflow(
      nodes = list(
          list(id = "start", type = "input", data = list(label = "Start Node"),
               position = list(x = 0, y = 0)),
          list(id = "mid", data = list(label = "Mid Node"),
               position = list(x = 0, y = 100)),
          list(id = "end", type = "output", data = list(label = "End Node"),
               position = list(x = 0, y = 200))
        )
      ,
      edges = list(
        list(source = "start", target = "mid", id = "start-mid"),
        list(source = "mid", target = "end", id = "mid-end")
      )
      
    )
  })
}

shinyApp(ui, server)
```
