---
title: "Fully Customized HTML Nodes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{fully_customized_html_nodes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
```

Following the tailwind example from reactflow [here](https://reactflow.dev/examples/styling/tailwind).
This is how we can have fully customized HTML nodes.

The basics of HTML nodes can be found in the Examples vignette.

In short, we have to set the `type` of each node to `"html"` and then we apply custom CSS to style each element.


## Global CSS styles

We will apply a mix of global CSS and direct styles (to stay in the spirit of Tailwind... :)).

Our global CSS will apply a general styling to each HTML node's border, size, radius, etc.
This is mostly copied from the default style for `react-flow__node-default`.
Then, we also style our handle a little-bit to make it a bit more distinct.

```{css}
.react-flow__node.react-flow__node-html {
  padding: 10px;
  border-radius: 15px;
  border: 4px solid #00808073;
  width: 150px;
  font-size: 12px;
  color: var(--xy-node-color, var(--xy-node-color-default));
  text-align: center;
  background-color: var(--xy-node-background-color, var(--xy-node-background-color-default));
}

.react-flow__handle{
  width: 3em;
  height: 5px;
  background-color: teal;
  border: none;
  border-radius: 0;
}
```


## Node and Edges Definition

Each node will have an icon, a name and title for three fictional employees.

```{r}
nodes <- list(
  list(id = "jd",
       data = list(icon = "bank2", name = "Jane Doe", title = "CEO"),
       position = list(x = 100, y = 0)),
  list(id = "tw",
       data = list(icon = "brush-fill", name = "Tyler Weary", title = "Designer"),
       position = list(x = 0, y = 100)),
  list(id = "kp", data = list(icon = "code-slash", name = "Kristi Price", title = "Developer"),
       position = list(x = 200, y = 100))
)

edges <- list(
  list(source = "jd", target = "tw", id = "js-tw"),
  list(source = "jd", target = "kp", id = "js-kp")
)
```

Note that we have not yet specified that the nodes are of `type = "html"`, neither have we created HTML code somwhere.

To abstract this from the node definitions, we will do this next in a separate function.

```{r}
# take a node element el and style it, using data$text, data$subtext, as well as data$icon
style_node <- function(el) {
  # 1. set the node type to html
  el$type <- "html"
  
  # 2. create the HTML code for each element but use as character
  el$data$html <- as.character(
    htmltools::div(
      style = paste(
        "display: flex; flex-direction: row; align-items: center;",
        "justify-content: space-around;"
      ),
      
      htmltools::div(
        style = paste(
          "background-color: #d1d1d187; border-radius: 10000px;",
          "height: 3em; width: 3em; display: flex; align-items: center;",
          "justify-content: center;"
        ),
        bsicons::bs_icon(el$data$icon, fill = "#00808080", 20)
      ),
      
      htmltools::div(
        htmltools::span(
          el$data$name,
          style = "font-weight: 800; color: teal;"
        ),
        htmltools::div(
          style = "font-size: 0.8em; font-weight: 600; color: #999999;",
          el$data$title
        )
      )
    )
  )
  el
}
```

Next, we will apply this styling to all nodes.

```{r}
nodes <- lapply(nodes, style_node)
```


## Draw the Graph

Last but not least, we have to draw the full flow:

```{r setup}
library(reactflow)
reactflow(
  nodes = nodes,
  edges = edges,
  mini_map(),
  background(),
  controls()
)
```
